---
title: "Wrangling Project"
output: html_notebook
---

```{r}
library(dplyr)
library(jsonlite)
library(knitr)
library(lubridate)
library(readr)
library(stringi)
library(stringr)
library(tibble)
library(tidyr)
```

## Data Gathering

```{r}
archive <- read.csv("twitter-archive-enhanced.csv", stringsAsFactors = FALSE)
```

```{r}
url <- "https://d17h27t6h515a5.cloudfront.net/topher/2017/August/599fd2ad_image-predictions/image-predictions.tsv"

img_predictions <- read.delim(url, sep = "\t", header = TRUE)
```

```{r}
api_data <- jsonlite::stream_in(file("tweet_json.txt"))
```

```{r}
api_data <- api_data %>%
  select(id, retweet_count, favorite_count)
```

## Assessing

#### Archive Data

```{r}
head(archive)
```

```{r}
kable(table(archive$source), col.names = c("Source", "Count"))
```

```{r}
head(archive$text)
```

```{r}
str(archive)
```

```{r}
kable(colSums(!is.na(archive)), col.names = c("Column", "Non-Null Count"))
```

```{r}
len <- nrow(table(archive$rating_numerator))

print(table(archive$rating_numerator)[1:(len/2)])
print(table(archive$rating_numerator)[(len/2):len])
```

```{r}
archive %>%
  filter(rating_numerator > 14) %>%
  select(text, rating_numerator, rating_denominator) %>%
  slice_head(n = 10)
```

📝 The 84/70 rating is for a group of dogs. This warrants further analysis...

```{r}
table(archive$rating_denominator)
```

#### Image Prediction Data

```{r}
head(img_predictions)
```

```{r}
str(img_predictions)
```

#### API Data

```{r}
head(api_data)
```

```{r}
str(api_data)
```

### Quality Issues

#### Archive Data

1.  Includes retweets. Per the `retweeted_` columns, there are 181 tweets.

2.  Includes tweets without URLs. There are 2356 entries, but only 2297 of them have `expanded_url`s.

3.  The `source` data includes HTML tags.

4.  The `text` data includes URLs referring back to the tweet itself. These URLs are shortened forms of those found in the `extended_urls` column.

5.  Some of the `expanded_urls` values consist of duplicate URLs strung together, separated by commas.

6.  The `rating_numerator` column

    -   lists only the fractional part of decimal numbers,
    -   lists the first numerator found in the `text` of each tweet---whether it's the rating or not.

7.  The `rating_denominator` column likewise lists the first denominator found in the `text` of each tweet---whether it's the rating or not.

8.  The `id` columns should be object-type, not float or int. And the `timestamp` column should be datetime.

### Tidiness Issues

1.  The tweet data spans multiple tables. The **`api_data`** consists of additional data on the tweets in **`archive`**, and the **`img_predictions`** are largely for the images in those tweets. As all three tables pertain to the same crop of tweets, they should be merged into a super table.

2.  The "dog stage" data spans multiple columns: `doggo`, `floofer`, `pupper`, and `puppo` are values of the "dog stage" variable.

## Cleaning Data

```{r}
# Make copies of original pieces of data
archive_clean <- archive
api_data_clean <- api_data
img_predict_clean <- img_predictions
```

### Issue 1: The 'archive' table includes retweets.

#### Define: Filter for tweets without `retweeted_` data. And deselect the now useless `retweeted_` columns.

\*If a tweet has `retweeted_` data, it's a retweet.

#### Code

```{r}
archive_clean <- archive_clean %>%
  filter(is.na(retweeted_status_id))
```

```{r}
archive_clean <- archive_clean %>%
  select(-retweeted_status_id, -retweeted_status_user_id,
         -retweeted_status_timestamp)
```

#### Test

```{r}
kable(names(archive_clean), col.names = "Column Names")
```

### Issue 2: The 'archive' table includes tweets without URLs.

#### Define: Filter for tweets with `expanded_urls`.

#### Code

```{r}
archive_clean <- archive_clean %>%
  filter(!is.na(expanded_urls))
```

#### Test

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

### Issue 3: The tweet data spans multiple tables.

#### Define: Join the tables, renaming the key columns as needed.

#### Code

```{r}
api_data_clean <- api_data_clean %>%
  rename(tweet_id = id)
```

```{r}
archive_clean <- inner_join(api_data_clean, archive_clean, by = "tweet_id") %>%
  inner_join(img_predict_clean, by = "tweet_id")
```

#### Test

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

### Issue 4: The "dog stage" data spans multiple columns: `doggo`, `floofer`, `pupper`, and `puppo` are values of the "dog stage" variable.

Most of the values are null.

#### Define: Unite the "dog stage" column's values into a new `dog_stage` column, but only the non-null values.

#### Code

```{r}
dog_stage_col <- c("doggo", "pupper", "floofer", "puppo")

archive_clean <- archive_clean %>%
  mutate(across(all_of(dog_stage_col), ~ ifelse(. == "None", NA_character_, .)))
```

```{r}
archive_clean <- archive_clean %>%
  unite(dog_stage, all_of(dog_stage_col), sep = "/", na.rm = TRUE)
```

#### Test

```{r}
archive_clean %>%
  slice(10:14) %>%
  select(tweet_id, dog_stage)
```

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

```{r}
unique(archive_clean$dog_stage)
```

### Issue 4.1: Empty strings are interpreted as non-null.

#### Define: Reassign the empty strings with None.

#### Code

```{r}
archive_clean$dog_stage <- ifelse(archive_clean$dog_stage == "", NA,
                                  archive_clean$dog_stage)
```

#### Test

```{r}
archive_clean %>%
  slice(10:14) %>%
  select(tweet_id, dog_stage)
```

```{r}
sum(!is.na(archive_clean$dog_stage))
```

```{r}
unique(archive_clean$dog_stage)
```

### Issue 5: The `source` data includes HTML tags.

#### Define: Replace the messy source data with just the part between the tags.

#### Code

```{r}
archive_clean$source <- str_replace_all(archive_clean$source,
                                        ".*>(.+)<.*", "\\1")
```

#### Test

```{r}
unique(archive_clean$source)
```

### Issue 6: The `text` data includes URLs referring back to the tweet itself.

#### Define: Remove the tweet URLs.

#### Code

```{r}
archive_clean$text <- sub(" https://t.co/[a-zA-Z0-9]+", "", archive_clean$text)
```

#### Test

```{r}
sum(grepl("https://t.co", archive_clean$text))
```

```{r}
head(archive_clean$text[grepl("https://t.co", archive_clean$text)])
```

✅ These URLs---separated by newlines---point to fundraiser pages: they are 
part of the tweet.

### Issue 7: Some of the `expanded_urls` values consist of duplicate URLs strung together, separated by commas.

#### Define: Split the multi-URL strings into URL lists. Then, extract the unique, list elements...

#### Code

```{r}
archive_clean$expanded_urls <- strsplit(archive_clean$expanded_urls, ",")

# Remove duplicates within each list element
archive_clean$expanded_urls <- lapply(
  archive_clean$expanded_urls, function(urls) {
    unique_urls <- unique(urls)
    as.character(unique_urls)
  }
)
```

#### Test

```{r}
multi_url <- archive_clean %>%
  filter(lengths(expanded_urls) > 1) %>%
  pull(expanded_urls)

head(multi_url)
tail(multi_url)
```

⚠️ While not duplicates per se, these fundraiser URLs are somewhat redundant 
with their shortened versions in the `test` data (above).

### Issue 7.1: Lists as elements do not make for tidy data frames!

Data should be discreet.

#### Define: Going down the list of expanded URL lists, for each sub-list:

-   reverse the order of elements such that the Twitter URL is always first,
-   paste the elements together into a string separated by commas,
-   and split the elements into separate columns.

#### Code

```{r}
archive_clean$expanded_urls <- lapply(
  archive_clean$expanded_urls, function(x) paste(rev(x), collapse = ", ")
)

```

```{r}
# The embedded URL column will be filled with `NA` where applicable.
archive_clean <- separate(
  archive_clean, expanded_urls,
  into = c("expanded_url_twitter", "expanded_url_embedded"),
  sep = ", "
)
```

### Test

```{r}
head(archive_clean)
```

```{r}
head(subset(archive_clean, select = c("expanded_url_twitter",
                                      "expanded_url_embedded")), 30)
```

### Issue 8: The `rating_numerator` column

-   lists only the fractional part of decimal numbers,
-   lists the first numerator found in the `text` of each tweet---whether it's the rating or not.

#### Define: Replace the `rating_numerator` values with complete rating numerators---integers and all---extracted from the `text` data.

#### Code

```{r}
# Find all sequences resembling a numerator and select the last the last one.
# The ratings are at the end.
archive_clean$rating_numerator <- sapply(
  str_extract_all(archive_clean$text, "\\d*\\.?\\d+/"),
  function(x) tail(x, n = 1)
)

archive_clean$rating_numerator <- str_replace(archive_clean$rating_numerator,
                                              "/$", "")
```

#### Test

```{r}
archive_clean %>%
  filter(str_detect(text, "\\d+\\.\\d+/")) %>%
  select(text, rating_numerator)
```

```{r}
archive_clean %>%
  select(text, rating_numerator)
```

```{r}
unique(archive_clean$rating_numerator)
```

```{r}
archive_clean %>%
  filter(rating_numerator == ".10") %>%
  select(text, rating_numerator)
```

### Issue 8.1: Periods immediately preceding the rating numerators were picked up by the RegEx.

#### Define: Strip preceding periods from the `rating_numerator` values.

#### Code

```{r}
archive_clean$rating_numerator <- as.numeric(
  sub("^\\.", "", archive_clean$rating_numerator)
)
```

#### Test

```{r}
unique(archive_clean$rating_numerator)
```

### Issue 9: The `rating_denominator` column lists the first denominator\* encountered in the `text` of each tweet.

Conversely, the rating denominators come at the end.

#### Define: Extract the last denominator found in each tweet's `text`.

#### Code

```{r}
archive_clean$rating_denominator <- stri_extract_last_regex(
  archive_clean$text, "\\/\\d+[\\s.,!]+|\\/\\d+$"
)

archive_clean$rating_denominator <- gsub(
  "[/. ,!]", "", archive_clean$rating_denominator
)
```

#### Test

```{r}
archive_clean %>%
  filter(is.na(rating_denominator)) %>%
  select(text, rating_numerator) %>%
  slice_head(n = 5)
```

```{r}
unique(archive_clean$rating_denominator)
```

### Issue 9.1: The RegEx pattern

-   isn't picking up denominators that are immediately followed by 's' or ')' characters,
-   is grabbing myriad, excess characters along with some of the denominators.

#### Define: Add 's' and')' to the RegEx, and drop the newline escape sequences.

#### Code

```{r}
archive_clean$rating_denominator <- stri_extract_last_regex(
  archive_clean$text, "\\/\\d+[\\s.,!)]+|\\/\\d+s\\s|\\/\\d+$"
)

archive_clean$rating_denominator <- as.numeric(
  gsub("[/. ,!)s\n]", "", archive_clean$rating_denominator)
)
```

#### Test

```{r}
unique(archive_clean$rating_denominator)
```

```{r}
archive_clean %>%
  filter(rating_denominator != "10") %>%
  select(text, rating_denominator)
```

### Issue 10: The `id` columns should be object-type, not float or int. And the `timestamp` column should be datetime.

#### Define: Convert the `id` columns to character-type and the `timestamp` column to datetime.

#### Code

```{r}
id_col_type <- c(
  tweet_id = "character",
  in_reply_to_status_id = "character",
  in_reply_to_user_id = "character"
)

archive_clean <- archive_clean %>%
  mutate(across(all_of(names(id_col_type)), as.character))

archive_clean$timestamp <- as_datetime(archive_clean$timestamp)
```

#### Test

```{r}
str(archive_clean)
```

## Storing Data

```{r}
write_csv(archive_clean, "twitter_archive_masteR.csv")
```
