---
title: "Wrangling Project"
output: html_notebook
---

```{r}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(knitr)
library(lubridate)
library(readr)
library(scales)
library(stringi)
library(stringr)
library(tibble)
library(tidyr)
```

## Data Gathering

```{r}
archive <- read.csv("twitter-archive-enhanced.csv", stringsAsFactors = FALSE)
```

```{r}
url <- "https://d17h27t6h515a5.cloudfront.net/topher/2017/August/599fd2ad_image-predictions/image-predictions.tsv"

img_predictions <- read.delim(url, sep = "\t", header = TRUE)
```

```{r}
api_data <- jsonlite::stream_in(file("tweet_json.txt"))
```

```{r}
api_data <- api_data %>%
  select(id, retweet_count, favorite_count)
```

## Assessing

#### Archive Data

```{r}
head(archive)
```

```{r}
kable(table(archive$source), col.names = c("Source", "Count"))
```

```{r}
head(archive$text)
```

```{r}
str(archive)
```

```{r}
kable(colSums(!is.na(archive)), col.names = c("Column", "Non-Null Count"))
```

```{r}
len <- nrow(table(archive$rating_numerator))

print(table(archive$rating_numerator)[1:(len/2)])
print(table(archive$rating_numerator)[(len/2):len])
```

```{r}
archive %>%
  filter(rating_numerator > 14) %>%
  select(text, rating_numerator, rating_denominator) %>%
  slice_head(n = 10)
```

📝 The 84/70 rating is for a group of dogs. This warrants further analysis...

```{r}
table(archive$rating_denominator)
```

#### Image Prediction Data

```{r}
head(img_predictions)
```

```{r}
str(img_predictions)
```

#### API Data

```{r}
head(api_data)
```

```{r}
str(api_data)
```

### Quality Issues

#### Archive Data

1.  Includes retweets. Per the `retweeted_` columns, there are 181 tweets.

2.  Includes tweets without URLs. There are 2356 entries, but only 2297 of them have `expanded_url`s.

3.  The `source` data includes HTML tags.

4.  The `text` data includes URLs referring back to the tweet itself. These URLs are shortened forms of those found in the `extended_urls` column.

5.  Some of the `expanded_urls` values consist of duplicate URLs strung together, separated by commas.

6.  The `rating_numerator` column

    -   lists only the fractional part of decimal numbers,
    -   lists the first numerator found in the `text` of each tweet---whether it's the rating or not.

7.  The `rating_denominator` column likewise lists the first denominator found in the `text` of each tweet---whether it's the rating or not.

8.  The `id` columns should be object-type, not float or int. And the `timestamp` column should be datetime.

### Tidiness Issues

1.  The tweet data spans multiple tables. The **`api_data`** consists of additional data on the tweets in **`archive`**, and the **`img_predictions`** are largely for the images in those tweets. As all three tables pertain to the same crop of tweets, they should be merged into a super table.

2.  The "dog stage" data spans multiple columns: `doggo`, `floofer`, `pupper`, and `puppo` are values of the "dog stage" variable.

## Cleaning Data

```{r}
# Make copies of original pieces of data
archive_clean <- archive
api_data_clean <- api_data
img_predict_clean <- img_predictions
```

### Issue 1: The 'archive' table includes retweets.

#### Define: Filter for tweets without `retweeted_` data. And deselect the now useless `retweeted_` columns.

\*If a tweet has `retweeted_` data, it's a retweet.

#### Code

```{r}
archive_clean <- archive_clean %>%
  filter(is.na(retweeted_status_id))
```

```{r}
archive_clean <- archive_clean %>%
  select(-retweeted_status_id, -retweeted_status_user_id,
         -retweeted_status_timestamp)
```

#### Test

```{r}
kable(names(archive_clean), col.names = "Column Names")
```

### Issue 2: The 'archive' table includes tweets without URLs.

#### Define: Filter for tweets with `expanded_urls`.

#### Code

```{r}
archive_clean <- archive_clean %>%
  filter(!is.na(expanded_urls))
```

#### Test

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

### Issue 3: The tweet data spans multiple tables.

#### Define: Join the tables, renaming the key columns as needed.

#### Code

```{r}
api_data_clean <- api_data_clean %>%
  rename(tweet_id = id)
```

```{r}
archive_clean <- inner_join(api_data_clean, archive_clean, by = "tweet_id") %>%
  inner_join(img_predict_clean, by = "tweet_id")
```

#### Test

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

### Issue 4: The "dog stage" data spans multiple columns: `doggo`, `floofer`, `pupper`, and `puppo` are values of the "dog stage" variable.

Most of the values are null.

#### Define: Unite the "dog stage" column's values into a new `dog_stage` column, but only the non-null values.

#### Code

```{r}
dog_stage_col <- c("doggo", "pupper", "floofer", "puppo")

archive_clean <- archive_clean %>%
  mutate(across(all_of(dog_stage_col), ~ ifelse(. == "None", NA_character_, .)))
```

```{r}
archive_clean <- archive_clean %>%
  unite(dog_stage, all_of(dog_stage_col), sep = "/", na.rm = TRUE)
```

#### Test

```{r}
archive_clean %>%
  slice(10:14) %>%
  select(tweet_id, dog_stage)
```

```{r}
kable(colSums(!is.na(archive_clean)), col.names = c("Column", "Non-Null Count"))
```

```{r}
unique(archive_clean$dog_stage)
```

### Issue 4.1: Empty strings are interpreted as non-null.

#### Define: Reassign the empty strings with None.

#### Code

```{r}
archive_clean$dog_stage <- ifelse(archive_clean$dog_stage == "", NA,
                                  archive_clean$dog_stage)
```

#### Test

```{r}
archive_clean %>%
  slice(10:14) %>%
  select(tweet_id, dog_stage)
```

```{r}
sum(!is.na(archive_clean$dog_stage))
```

```{r}
unique(archive_clean$dog_stage)
```

### Issue 5: The `source` data includes HTML tags.

#### Define: Replace the messy source data with just the part between the tags.

#### Code

```{r}
archive_clean$source <- str_replace_all(archive_clean$source,
                                        ".*>(.+)<.*", "\\1")
```

#### Test

```{r}
unique(archive_clean$source)
```

### Issue 6: The `text` data includes URLs referring back to the tweet itself.

#### Define: Remove the tweet URLs.

#### Code

```{r}
archive_clean$text <- sub(" https://t.co/[a-zA-Z0-9]+", "", archive_clean$text)
```

#### Test

```{r}
sum(grepl("https://t.co", archive_clean$text))
```

```{r}
head(archive_clean$text[grepl("https://t.co", archive_clean$text)])
```

✅ These URLs---separated by newlines---point to fundraiser pages: they are 
part of the tweet.

### Issue 7: Some of the `expanded_urls` values consist of duplicate URLs strung together, separated by commas.

#### Define: Split the multi-URL strings into URL lists. Then, extract the unique, list elements...

#### Code

```{r}
archive_clean$expanded_urls <- strsplit(archive_clean$expanded_urls, ",")

# Remove duplicates within each list element
archive_clean$expanded_urls <- lapply(
  archive_clean$expanded_urls, function(urls) {
    unique_urls <- unique(urls)
    as.character(unique_urls)
  }
)
```

#### Test

```{r}
multi_url <- archive_clean %>%
  filter(lengths(expanded_urls) > 1) %>%
  pull(expanded_urls)

head(multi_url)
tail(multi_url)
```

⚠️ While not duplicates per se, these fundraiser URLs are somewhat redundant 
with their shortened versions in the `test` data (above).

### Issue 7.1: Lists as elements do not make for tidy data frames!

Data should be discreet.

#### Define: Going down the list of expanded URL lists, for each sub-list:

-   reverse the order of elements such that the Twitter URL is always first,
-   paste the elements together into a string separated by commas,
-   and split the elements into separate columns.

#### Code

```{r}
archive_clean$expanded_urls <- lapply(
  archive_clean$expanded_urls, function(x) paste(rev(x), collapse = ", ")
)

```

```{r}
# The embedded URL column will be filled with `NA` where applicable.
archive_clean <- separate(
  archive_clean, expanded_urls,
  into = c("expanded_url_twitter", "expanded_url_embedded"),
  sep = ", "
)
```

### Test

```{r}
head(archive_clean)
```

```{r}
head(subset(archive_clean, select = c("expanded_url_twitter",
                                      "expanded_url_embedded")), 30)
```

### Issue 8: The `rating_numerator` column

-   lists only the fractional part of decimal numbers,
-   lists the first numerator found in the `text` of each tweet---whether it's
    the rating or not.

#### Define: Replace the `rating_numerator` values with complete rating numerators---integers and all---extracted from the `text` data.

#### Code

```{r}
# Find all sequences resembling a numerator and select the last the last one.
# The ratings are at the end.
archive_clean$rating_numerator <- sapply(
  str_extract_all(archive_clean$text, "\\d*\\.?\\d+/"),
  function(x) tail(x, n = 1)
)

archive_clean$rating_numerator <- str_replace(archive_clean$rating_numerator,
                                              "/$", "")
```

#### Test

```{r}
archive_clean %>%
  filter(str_detect(text, "\\d+\\.\\d+/")) %>%
  select(text, rating_numerator)
```

```{r}
archive_clean %>%
  select(text, rating_numerator)
```

```{r}
unique(archive_clean$rating_numerator)
```

```{r}
archive_clean %>%
  filter(rating_numerator == ".10") %>%
  select(text, rating_numerator)
```

### Issue 8.1: Periods immediately preceding the rating numerators were picked up by the RegEx.

#### Define: Strip preceding periods from the `rating_numerator` values.

#### Code

```{r}
archive_clean$rating_numerator <- as.numeric(
  sub("^\\.", "", archive_clean$rating_numerator)
)
```

#### Test

```{r}
unique(archive_clean$rating_numerator)
```

### Issue 9: The `rating_denominator` column lists the first denominator\* encountered in the `text` of each tweet.

Conversely, the rating denominators come at the end.

#### Define: Extract the last denominator found in each tweet's `text`.

#### Code

```{r}
archive_clean$rating_denominator <- stri_extract_last_regex(
  archive_clean$text, "\\/\\d+[\\s.,!]+|\\/\\d+$"
)

archive_clean$rating_denominator <- gsub(
  "[/. ,!]", "", archive_clean$rating_denominator
)
```

#### Test

```{r}
archive_clean %>%
  filter(is.na(rating_denominator)) %>%
  select(text, rating_denominator) %>%
  slice_head(n = 5)
```

```{r}
unique(archive_clean$rating_denominator)
```

### Issue 9.1: The RegEx pattern

-   isn't picking up denominators that are immediately followed by 's' or ')' characters,
-   is grabbing myriad, excess characters along with some of the denominators.

#### Define: Add 's' and')' to the RegEx, and drop the newline escape sequences.

#### Code

```{r}
archive_clean$rating_denominator <- stri_extract_last_regex(
  archive_clean$text, "\\/\\d+[\\s.,!)]+|\\/\\d+s\\s|\\/\\d+$"
)

archive_clean$rating_denominator <- as.numeric(
  gsub("[/. ,!)s\n]", "", archive_clean$rating_denominator)
)
```

#### Test

```{r}
unique(archive_clean$rating_denominator)
```

```{r}
archive_clean %>%
  filter(rating_denominator != "10") %>%
  select(text, rating_denominator)
```

### Issue 10: The `id` columns should be object-type, not float or int. And the `timestamp` column should be datetime.

#### Define: Convert the `id` columns to character-type and the `timestamp` column to datetime.

#### Code

```{r}
id_col_type <- c(
  tweet_id = "character",
  in_reply_to_status_id = "character",
  in_reply_to_user_id = "character"
)

archive_clean <- archive_clean %>%
  mutate(across(all_of(names(id_col_type)), as.character))

archive_clean$timestamp <- as_datetime(archive_clean$timestamp)
```

#### Test

```{r}
str(archive_clean)
```

## Storing Data

```{r}
write_csv(archive_clean, "twitter_archive_masteR.csv")
```

```{r}
df <- read.csv("twitter_archive_masteR.csv")
```

```{r}
str(df)
```

```{r}
df$tweet_id <- as.character(df$tweet_id)
```

## Analyzing and Visualizing Data

```{r}
rating_cols = c("rating_numerator", "rating_denominator", "text", "jpg_url")

subset(df, rating_denominator > 10, select = rating_cols)
```

### Insight #1: The `rating_denominator` is $10 \times n$, where $n$ is the number of dogs.

-   There's normally one dog, and the normal denominator is 10.
-   #988 has five dogs and a denominator of 50.
-   #946 has nine puppies and a rating of 99 out of 90---an average rating of 11/10.

```{r}
# Normalization of ratings
df$normalized_rating <- (df$rating_numerator / df$rating_denominator) * 10

summary(df$normalized_rating)
```

```{r}
boxplot_stats_result <- boxplot.stats(df$normalized_rating)
lower_limit <- boxplot_stats_result$stats[1]
upper_limit <- boxplot_stats_result$stats[5]
```

```{r}
df_sans_outlier <- subset(df, normalized_rating >= lower_limit & normalized_rating <= upper_limit)
```

```{r}
summary(df_sans_outlier$normalized_rating)
```

```{r}
mean(df$rating_numerator)
```

```{r}
df_outliers <- subset(df, normalized_rating < lower_limit | normalized_rating > upper_limit)

subset(df_outliers, rating_numerator > upper_limit, select = rating_cols)
```

### Insight #1.1: Normalized and stripped of outliers, the mean rating is 10.9 out of 10.

-   With outliers, the mean is 11.7.
    -   Largely due to a couple of gag tweets (above).
-   Prior to normalization, the mean rating was 12.2.

```{r}
conf_cols <- c("p1_conf", "p2_conf", "p3_conf")
subset_df <- subset(df_sans_outlier, select = conf_cols)

summary(subset_df)
```

### Insight #2: The neural network is not very confident in its dog breed predictions.

-   The median confidence level is 0.60, and that's for its best guesses!
-   The five-number summary is:

```         
    - minimum      4%
    - 1st quartile 37%
    - median       60%
    - 3rd quartile 85%
    - maximum      100%
```

```{r}
# If we are to analyze the tweets on a per-breed basis, we need to be reasonably
# sure that a Pug is—in fact—a Pug
by_breed <- df_sans_outlier %>%
  filter(p1_conf >= 0.80 & p1_dog == "True")
```

```{r}
num_dogs <- by_breed %>%
  group_by(p1) %>%
  summarize(number_of_dogs = n())

head(num_dogs)
```

```{r}
mean_rating <- by_breed %>%
  group_by(p1) %>%
  summarize(mean_rating = mean(normalized_rating, na.rm = TRUE))

head(mean_rating)
```

```{r}
df_by_breed <- full_join(num_dogs, mean_rating)
df_by_breed_sort <- arrange(df_by_breed, mean_rating)
```

```{r}
head(df_by_breed_sort, 5)
```

```{r}
tail(df_by_breed_sort, 5)
```

```{r}
df_by_breed %>%
  group_by(number_of_dogs) %>%
  summarize(mean_rating = mean(mean_rating)) %>%
  arrange(number_of_dogs)

```

### Insight #3: Being unique appears to give a ratings boost.

-   The highest-rated dogs are all unique.
-   On average, the most popular breeds tend to be higher rated than the least common breeds:
    -   Unique breeds are the exception.

```{r}
by_breed_mean <- by_breed %>%
  group_by(p1) %>%
  summarize(
    mean_normalized_rating = mean(normalized_rating),
    mean_favorite_count = mean(favorite_count)
  )
```

```{r}
by_breed_mean
```

```{r}
ggplot(by_breed_mean, aes(x = mean_normalized_rating, y = mean_favorite_count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "#3274A1") +
  labs(
    x = "Normalized Rating",
    y = "Favorite Count",
    title = "Breed Rating vs Popular Preference"
  ) +
  scale_y_continuous(labels = comma_format(scale = 1e-3, suffix = "k"))
```

```{r}
r <- cor(by_breed_mean$mean_normalized_rating, by_breed_mean$mean_favorite_count)
R2 <- r^2

cat("      Correlation coefficient (r):    ", r, "\n")
cat("Coefficient of determination (R2):    ", R2)
```

### Insight #4: The ratings correlate quite well with popular preference.

-   There is a moderate, positive correlation between normalized rating and favorite count.
    -   The correlation coefficient ($r$) is 0.5982.
    -   The coefficient of determination ($R^2$) is 0.358.
-   Favorite count is a [reflection of user preference](https://medium.com/@Encore/favorites-vs-retweets-and-why-one-is-more-important-than-the-other-ba12ee20e9ba). The most-favorited dog breeds are the most preferred.

### Insights Recap:
1. The `rating_denominator` is $10 \times n$, where $n$ is the number of dogs.

    1.1. Normalized and stripped of outliers, the mean rating is 10.9 out of 10.

2. The neural network is not very confident in its dog breed predictions.

3. Being a unique breed appears to give a ratings boost.

4. The ratings correlate quite well with popular preference, though they are not very predictive.*

*This is normal for explanatory variables where the response variable is human behavior.